FraCaS: Temporal Analysis

### 2020-06-23 paper2020/paper.tex
revise intro

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the
years~\citep{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}. The
semantics of tense and aspect has been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7). This part
of the FraCaS test suite has been notoriously difficult to tackle. In
fact, all the computational systems ran on the FraCaS test suite so
far skip this section. Additionally, they also skip the anaphora
and ellipsis sections, with the exception of the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019}, which offers a complete whole run of the
FraCaS test suite, excluding the temporal section. In this
paper, we take up the challenge of providing a computationally viable
account of tense and aspect to deal with the notorious section 7 of
the FraCaS test suite. Our account is not meant to be a theoretically
full account of tense and aspect, but rather an account that is driven by the
need to cover the test suite in a way that is general enough to
capture the test suite examples. The account is evaluated on the
entailment properties of various temporal and aspectual examples, as
given by the test suite. As such, we are not getting into the
discussion of how tense and aspect might affect grammaticality or
infelicitousness of various sentences. We take the sentences of the
FraCaS suite to be syntactically correct and the entailment patterns they
are annotated with valid.

The paper is structured as follows: in \cref{sec:background}, we give
a brief summary of the frameworks whose various subsystems rely on. In
particular, the Grammatical Framework is used to construct the
syntactic parser; the Coq proof assistant checks all the reasoning; a
monad-based dynamic semantics deals with references (anaphora).  We
also provide some brief remarks on temporal semantics. In \cref{sec:our-semantics},
we discuss the main aspects of the compositional semantics of our, using various examples
from the FraCaS suite in exemplifying its effectiveness. In \cref{sec:results},
we take a brief look at how our system performs with respect to the
FraCaS suite. We ran the system across the whole suite, this being the
first time that a complete run through the whole suite has been made,
but we look in particular at the performance on the temporal
section. In \cref{sec:conclusions}, we conclude and discuss avenues for future
work.

%We do not care about classifying sentences as felicitous/infelicitous
%or grammatical/ungrammatical, rather we assume that the sentence is
%correct and care about giving it an appropriate interpretation as a
%logical formula.  Indeed, our evaluation uses entailment properties
%only.

>>>>

The semantics of tense and aspect has been a long standing issue in
the study of formal semantics since the early days of Montague Grammar
and a number of different ideas have been put forth on the table to
deal with them throughout the years. Recent proposals include the
works of the following authors:
\citet{dowty:2012,prior:2003,steedman:2000,higginbotham:2009,fernando:2015}.
%
The
semantics of tense and aspect have been also considered in the study of
Natural Language Inference (NLI). The various datasets
for NLI that have been proposed by the years contain examples that
have some implicit or explicit reliance on inferences related to tense
and aspect. One of the early datasets used to test logical approaches,
the FraCaS test suite~\citep{cooper:1996} contains a whole section
dedicated to temporal and aspectual inference (section 7 of the dataset). This part
of the FraCaS test suite has been notoriously difficult to tackle. In
fact, so far, no computational system has been capable to deal with it: when
authors report accuracy over the FraCaS test suite they
skip this section.
%
In fact, they also often skip the anaphora and ellipsis sections, the
exception being the system presented by
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019},
which include support for anaphora and ellipsis but still omit the
temporal section.
%
In this paper, we take up the challenge of providing a
computationally viable account of tense and aspect to deal with the
notoriously difficult section 7 of the FraCaS test suite. Our account is not meant
to be a theoretically extensive account of tense and aspect, but rather an
account that is driven by the need to cover the test suite in a way
that is general enough to capture the test suite examples, \emph{while still covering the rest of the FraCaS test suite}.
%
The account is evaluated on the entailment properties of various
temporal and aspectual examples, as given by the test suite. As such,
we are not getting into the discussion of how tense and aspect might
affect grammaticality or infelicitousness of various sentences.
We assume that the sentences of the FraCaS suite are syntactically and
semantically correct, and strive to produce accurate logical
representations given that assumption. We further assume that the
entailment annotations of various problems are valid, and we use those
to evaluate the correctness of the logical representations of
sentences.

The paper is structured as follows: in \cref{sec:background}, we give
a brief summary of the computational frameworks whose various
subsystems rely on. In particular, the Grammatical Framework is used
to construct the syntactic parser, the Coq proof assistant checks all
the reasoning and a monad-based dynamic semantics deals with
Montague-style semantics, and references (anaphora).  We also provide
some brief remarks on temporal semantics. In \cref{sec:our-semantics},
we discuss the main aspects of the compositional semantics of our
system, using various examples from the FraCaS suite to illustrate its
effectiveness. In \cref{sec:results}, we evaluate how our system
performs with respect to the FraCaS suite. We ran the system across
the whole suite: our system is thus the first which is capable of
handling the complete FraCaS test suite.  Yet, we are interested in
particular in the performance on the temporal section. In
\cref{sec:conclusions}, we conclude and discuss avenues for future
work.
